---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mortar <img src="man/figures/logo.svg" width="20%" align="right" alt="Hex logo for mortar package. Shows a partially filled water pipe coming out of a brick wall where there is generous amounts of mortar between bricks." />

<!-- badges: start -->

<!-- badges: end -->

<!-- > *Mortar: a plastic building material (such as a mixture of cement, -->
<!-- > lime, or gypsum plaster with sand and water) that hardens and is used -->
<!-- > in masonry or plastering* -->
<!-- > ([source](https://www.merriam-webster.com/dictionary/mortar)) -->

## Overview

> *"If you need to copy + paste the same code more than once, write a function. If you need to write the same function more than once, create a package."*
>
> –DRY coding principle

`mortar` was created to standardize common workflows in the USGS Data Science Community of Practice (DS CoP) to produce more robust, reproducible pipelines. It contains helper functions to standardize (1) the organization of project repositories and (2) the creation of [`targets`](https://books.ropensci.org/targets/) pipelines using the DS CoP best practices. We draw upon community developed best practices as well as certain USGS-specific requirements. We welcome feedback and suggestions from the DS CoP community and beyond!

#### Where did the name come from?
Thepackage name has multiple meanings. The package mixes water (😉) with structural ingredients to form a foundation upon which our pipelines can be more easily and efficiently built. It was also initially conceptualized to provide **mor**e **tar**gets functions than those available in the `targets` and `tarchetypes` packages.

## Installation

Install the package from source from [GitHub](https://github.com/DOI-USGS/mortar):

``` r
# Install the pak package if you don't already have it
#install.packages("pak")

pak::pak("DOI-USGS/mortar")
```

## Usage

The main two functions in this package are `use_project_usgs()` and `tar_init()`. These build the structure of your projects and targets pipelines, removing much of the tedium of starting up a project and/or pipeline.

### `use_project_usgs()`
Use `use_project_usgs()` to setup your project repository. It will create all the metadata files necessary for a software release, with options to customize the files. These files can then be manually edited throughout the development process to keep metadata up to date.

```{r eval=FALSE}
library(mortar)

use_project_usgs()
```

There are additional `use_*_usgs()` functions to add common project files to a directory such as a `.gitignore`, `README.md`, if you want to add them individually without setting up an entire project.

### `tar_init()`
If you're setting up a `targets` pipeline project, you can **also** run `tar_init()` to set up the `targets` directory structure we commonly use in the DS CoP. `tar_init()` generates the targets script file (`_targets.R`), split phase directories (optional; e.g., `01_fetch/`), and subdirectories (e.g., `02_process/src/`). 

```{r eval=FALSE}
tar_init(
  phase_names = c("fetch", "process", "visualize"),
  seperate_phase_scripts = TRUE,
  phase_subdirs = c("in", "src", "out"),
  use_leading_zeros = TRUE
)
```

Refer to the [Get started](https://doi-usgs.github.io/mortar/articles/mortar.html) page for more information using `mortar.`

## Consider contributing
Please consider reporting bugs, asking questions, or suggesting enhancements. This is a community developed package and we want it to be inclusive of the needs of our community. If considering contributing, please read [CONTRIBUTING.md](https://github.com/DOI-USGS/mortar/blob/main/CONTRIBUTING.md) for more information. Also, please read through [CONDUCT.md](https://github.com/DOI-USGS/mortar/blob/main/CONDUCT.md) to ensure that we maintain a warm, welcoming, safe, and professional environment.

## Disclaimer
This software is preliminary or provisional and is subject to revision. It is being provided to meet the need for timely best science. The software has not received final approval by the U.S. Geological Survey (USGS). No warranty, expressed or implied, is made by the USGS or the U.S. Government as to the functionality of the software and related material nor shall the fact of release constitute any such warranty. The software is provided on the condition that neither the USGS nor the U.S. Government shall be held liable for any damages resulting from the authorized or unauthorized use of the software.
