---
title: "mortar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mortar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of `mortar` is to standardize how we in the IIDD Data Science Branch structure our collaborative projects.
The package's functionality focuses on setting up project directories using our branch's conventions.
Think of `mortar` as a collection of helper functions for creating files and folders we often use in our workflows.

```{r setup}
library(mortar)
```

## Set up a `targets` project with `tar_init()`

The `tar_init()` function will initialize a `targets` pipeline project directory according to our branch's convention described in more detail [here](https://wma.code-pages.usgs.gov/dsp/trainings/ds-pipelines-targets-2-course/12-usgs-data-science-conventions.html).
Run this function at the beginning of a `targets` pipeline project to avoid manually setting up the project directory (and forgetting to add those pesky `.empty` files in each sub-directory).

The call below illustrates how to use `tar_init()`.

```{r eval=FALSE}
tar_init(phase_names = c("fetch","process","summarize"),
         phase_nums = 1:3,
         phase_subdirs = c("src","out"))
```

Under the hood, `tar_init()` basically just repeatedly calls `dir.create()` and `file.create()` to accomplish the following tasks:

1. Creates a `_targets.R` file in the project's root directory with some general starter code

2. Creates `.R` scripts and directories  in the projects root directory for each phase of the pipeline like `1_fetch.R` and `1_fetch/`, etc.

3. For each phase directory, creates sub-directories therein such as `src` and `out`

4. Adds a `.empty` file within each sub-directory so that it will be tracked by git

Run `?mortar::tar_init` to view the function's documentation and more examples.

## Set up a USGS git repo with `use_*_usgs()`

You may want/need to include other common files in a project's git repo to adhere to coding conventions or standards.
These include a **README** that summarizes basic information about a repo, a **License** that details how the contents of a repo can be used, and others (see [here](https://code.usgs.gov/wma/national-iwaas/NWAA/standard-template-repository/) for more details).

The `use_*_usgs()` family of functions create these sorts of files based on USGS-specific templates available [here](https://code.usgs.gov/wma/national-iwaas/NWAA/standard-template-repository).
These functions are heavily inspired by similar functions from the [`usethis`](https://usethis.r-lib.org/) R package.

For example, you can add a .gitignore file to your repo with some pre-set (often forgotten) ignores using the following.
If you don't like the contents of these template files, you're welcome to edit them afterward to your needs.

```{r eval=FALSE}
use_gitignore_usgs()
```


